---
title: "R Notebook"
output: html_notebook
---

Operaciones de descripción y agrupación sobre tablas

Operaciones agregantes
apply y sapply aplican una función agregante por fila ó columna


The Apply family comprises: apply, lapply , sapply, vapply, mapply, rapply, and tapply. The Family of Apply functions pertains to the R base package, and is populated with functions to manipulate slices of data from matrices, arrays, lists and data frames in a repetitive way. 

Apply y el resto de funciones derivadas, están diseñadas para evitar utilizar bucles durante el tratamiento de datos. Actúan tomando como entrada una matriz, una lista o un array y aplican una función.

La función Apply, devuelve un vector, matriz o lista de valores, obtenidos al aplicar una función a una matriz o submatriz.
```{r}
A <- array(1:20,dim=c(4,5))
A
help(apply) #Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.
apply(A,1,sum)  #En el segundo parámetro 1 indica filas, 2 columnas
x<-apply(A,2,sum)  
x
class(x)
x<-apply(A,2,prod) 
x
class(x)
x<-apply(A,2,max)
x
class(x)
x<-apply(A,2,min)
x
class(x)
apply(A,1,is.vector)
apply(A,2,is.vector)


```

Ejemplos con un set de datos concreto
```{r}
data(cars)
help(cars) #descripción del dataset
dim(cars)
str(cars)
head(cars)
```

Resumen del dataset
```{r}
summary(cars)
```

Estadísticos de agregación
```{r}
apply(cars,2,min)  #valor mínimo para velocidad y distancia
apply(cars,2,max)  #valor mínimo para velocidad y distancia
apply(cars,2,mean)  #valor mínimo para velocidad y distancia
cars
```

lapply
You want to apply a given function to every element of a list and obtain a list as result. When you execute ?lapply

The difference is that:

It can be used for other objects like dataframes, lists or vectors; and
The output returned is a list (which explains the “l” in the function name), which has the same number of elements as the object passed to it.

```{r}
#lapply returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X.
lapply(cars,function(cars) cars/2) #lapply devuelve una lista como resultado
#the above lapply function divides the values in the dataframe by 2 and the
#output will be in form of list

lista_1<-lapply(cars,mean) #devuelve el resultado en forma de lista
lista_1$speed
```

sapply function takes list, vector or Data frame  as input. It is similar to lapply function but returns only vector as output
```{r}
sapply(cars,class)
vector_s<-sapply(cars,function(cars) cars/2) #sapply devuelve un vector como resultado
vector_s[4:10]
```

```{r}
sapply(cars,log) #devuelve un vector como resultado, aplicando la función a cada elemento
apply (cars,1,log) 
```

mapply: es una versión multivariable de sapply. mapply aplica FUN a los primeros elementos de cada argumento (...), los segundos elementos, los terceros elementos, etc.
```{r}
mapply(sum, 1:4)
mapply(sum, 1:4,1:4)
mapply(sum, 1:4,1:4,1:4)
#mapply sums up all the first elements(1+1+1) ,sums up all the
#second elements(2+2+2) and so on so the result will be


mapply(mean, cars)
```



tapply Para cuando desea aplicar una función a los subconjuntos de un vector y los subconjuntos están definidos por algún otro vector, generalmente un factor.
```{r}
help(tapply)
hijos<-c(1,3,2,2,3,5,3,2,1,2,3,1,2,4,1,0)
prov<-c("lu","co","lu","po","co","or","co","lu","lu","po","co","or","co","co","lu","po")
provf<-factor(prov)
tapply(hijos,provf,mean) #hace la media por categorías relacionando las listas hijoas y provf

```

